The Evolution of Phenotypically Invariant Genetic Networks

### 2017-12-01 kalman_walk_paper.tex
general edits to paper

\begin{scope}[>={Stealth[black]},
              every node/.style={fill=white,circle},
              every edge/.style={draw=black, thick}]
    \path [->, sloped] (A) edge[bend left] node {\tiny $2.2$} (B);
   % \path [->, sloped] (B) edge[bend left] node {\tiny ?} (A); 
    \path [->, sloped] (A) edge[bend right] node {\tiny $2$} (C); 
    %\path [->, sloped] (B) edge[bend right] node {\tiny ?} (C); 
    \path [->, >=Rectangle, sloped] (C) edge[bend right] node {\tiny $-1$} (A); 
    \path [->, >=Rectangle, sloped] (C) edge[bend right] node {\tiny $-2.2$} (B); 
    \path[->] (U) edge node {\tiny $1$} (A);
    \path[->] (U) edge node {\tiny $1$} (B);
    \path[->] (A) edge node {\tiny $1$} (y);
\end{scope}
\begin{scope}[>={Stealth[black]},
              every edge/.style={draw=black, thick}]
    \path [->] (A) edge[loop below] node[sloped, anchor=center, below] {\tiny $1$} (A);
    \path [->] (B) edge[loop above] node[sloped, anchor=center, above] {\tiny $1$} (B);
    \path [->, >=Rectangle] (C) edge[loop right] node[sloped, anchor=center, above] {\tiny $-1$} (C);
\end{scope}
\end{tikzpicture}
  \caption{A possible non-minimal three-gene oscillator phenotypically equivalent to the systems in Examples \ref{ex:oscillator} and \ref{ex:all_osc}.}
  \label{fig:3D_osc}
\end{figure}

@@

\paragraph{Hybrid incompatibility}
Two parents with the optimal phenotype can produce offspring whose phenotype is suboptimal
if the parents have different underlying systems.
%This leads to the question: 
How quickly do hybrid phenotypes break down as genetic distance between parents increases?
To quantify this, we will measure how far a system's phenotype is from optimal
using a weighted difference between impulse response functions.
Suppose that $\rho(t)$ is a nonnegative, smooth, square-integrable weighting function,
suppose that $h_0(t)$ is the \emph{optimal} impulse response function 
and define the ``distance to optimum'' of another impulse response function
to be
\begin{align}
\label{eqn:distance}
	D(h) = \left( \int_0^\infty \rho(t) \|h(t) - h_0(t)\|^2 dt \right)^{1/2} .
\end{align}
Consider reproduction between a parent with system $(A, B, C)$ 
and another displaced by distance $\epsilon$ in the direction $(X,Y,Z)$,
i.e., having  system $(A + \epsilon X, B + \epsilon Y, C + \epsilon Z)$.
We assume both are ``perfectly adapted'' systems, 
i.e., having impulse response function $h_0(t)$,
and their offspring has impulse response function $h_\epsilon(t)$.
A Taylor expansion of $D(h_\epsilon)$ in $\epsilon$ is explicitly worked out in Appendix \ref{apx:H_calc},
and shows that the phenotype of an $F_1$ hybrid between these two is at distance proportional to $\epsilon^2$ from optimal,
while $F_2$ hybrids are at distance proportional to $\epsilon$.
This is because an $F_1$ hybrid has one copy of each parental system,
and therefore lies directly between the parental systems (see Figure \ref{fig:conceptual_fig}) --
the parents both lie in $\allS$, which is the valley defined by $D$,
and so their midpoint only differs from optimal due to curvature of $\allS$.
In contrast, an $F_2$ hybrid may be homozygous for one parental type in some coefficients
and homozygous for the other parental type in others;
this means that each coefficient of an $F_2$ may be equal to either one of the parents,
or intermediate between the two;
this means that possible $F_2$ systems may be as far from the optimal set, $\allS$,
as the distance between the parents.
The precise rate at which the phenotype of a hybrid diverges depends on the geometry --
in Figure \ref{fig:conceptual_fig}, this is depicted as the angle of the black line (the optimal set)
with respect to the coordinates.

@@

Using simulations, \citet{porter2002speciation} demonstrated the accumulation of hybrid incompatibilities under directional, but not stabilizing selection, and \citet{palmer2009dynamics} observed the appearence of incompatibilities in suboptimally adapted populations in a constant environment. 

@@

Finally, despite our model's precise separation of phenotype and kryptotype, this relationship in nature may be far more complicated as aspects of the kryptotype may be less ``hidden'' than we currently assume. For instance, attributes excluded from the phenotype as modelled here, ignore the potential energy costs associated with execessively large (nonminimal) kryptotypes, as well as the relationship between a specific network architecture and mutational robustness, as well as many other presently neglected phenomena. Despite the model's simplifications and potentially nontrivial oversights, it may still roughly reflect and shed light upon speciation patterns as they occur in nature.

>>>>

\begin{scope}[>={Stealth[black]},
              every node/.style={fill=white,circle},
              every edge/.style={draw=black, thick}]
    \path [->, sloped] (A) edge[bend left] node {\tiny $2.2$} (B);
   % \path [->, sloped] (B) edge[bend left] node {\tiny ?} (A); 
    \path [->, sloped] (A) edge[bend right] node {\tiny $2$} (C); 
    %\path [->, sloped] (B) edge[bend right] node {\tiny ?} (C); 
    \path [->, >=Rectangle, sloped] (C) edge[bend right] node {\tiny $-1$} (A); 
    \path [->, >=Rectangle, sloped] (C) edge[bend right] node {\tiny $-2.2$} (B); 
    \path[->] (U) edge node {\tiny $1$} (A);
    \path[->] (U) edge node {\tiny $1$} (B);
    \path[->] (A) edge node {\tiny $1$} (y);
\end{scope}
\begin{scope}[>={Stealth[black]},
              every edge/.style={draw=black, thick}]
    \path [->] (A) edge[loop below] node[sloped, anchor=center, below] {\tiny $1$} (A);
    \path [->] (B) edge[loop above] node[sloped, anchor=center, above] {\tiny $1$} (B);
    \path [->, >=Rectangle] (C) edge[loop right] node[sloped, anchor=center, above] {\tiny $-1$} (C);
\end{scope}
\end{tikzpicture}
  \caption{A possible non-minimal three-gene oscillator, phenotypically equivalent to $A(\tau)$, the systems in Examples \ref{ex:oscillator} and \ref{ex:all_osc}.}
  \label{fig:3D_osc}
\end{figure}

@@

\subsection*{Hybrid incompatibility}
Two parents with the optimal phenotype can produce offspring whose phenotype is suboptimal
if the parents have different underlying systems.
%This leads to the question: 
How quickly do hybrid phenotypes break down as genetic distance between parents increases?
To quantify this, we will measure how far a system's phenotype is from optimal
using a weighted difference between impulse response functions.
Suppose that $\rho(t)$ is a nonnegative, smooth, square-integrable weighting function,
suppose that $h_0(t)$ is the \emph{optimal} impulse response function 
and define the ``distance to optimum'' of another impulse response function
to be
\begin{align}
\label{eqn:distance}
	D(h) = \left( \int_0^\infty \rho(t) \|h(t) - h_0(t)\|^2 dt \right)^{1/2} .
\end{align}
Consider reproduction between a parent with system $(A, B, C)$ 
and another displaced by distance $\epsilon$ in the direction $(X,Y,Z)$,
i.e., having  system $(A + \epsilon X, B + \epsilon Y, C + \epsilon Z)$.
We assume both are ``perfectly adapted'' systems, 
i.e., having impulse response function $h_0(t)$,
and their offspring has impulse response function $h_\epsilon(t)$.
A Taylor expansion of $D(h_\epsilon)$ in $\epsilon$ is explicitly worked out in Appendix \ref{apx:H_calc},
and shows that the phenotype of an $F_1$ hybrid between these two is at distance proportional to $\epsilon^2$ from optimal,
while $F_2$ hybrids are at distance proportional to $\epsilon$.
This is because an $F_1$ hybrid has one copy of each parental system,
and therefore lies directly between the parental systems (see Figure \ref{fig:conceptual_fig}) --
the parents both lie in $\allS$, which is the valley defined by $D$,
and so their midpoint only differs from optimal due to curvature of $\allS$.
In contrast, an $F_2$ hybrid may be homozygous for one parental type in some coefficients
and homozygous for the other parental type in others;
this means that each coefficient of an $F_2$ may be equal to either one of the parents,
or intermediate between the two;
this means that possible $F_2$ systems may be as far from the optimal set, $\allS$,
as the distance between the parents.
The precise rate at which the phenotype of a hybrid diverges depends on the geometry --
in Figure \ref{fig:conceptual_fig}, this is depicted as the angle of the black line (the optimal set)
with respect to the coordinates.

@@

Using simulations, \citet{porter2002speciation} demonstrated the accumulation of hybrid incompatibilities under directional, but not stabilizing selection, and \citet{palmer2009dynamics} observed the appearence of incompatibilities in populations evolving in a constant environment, only if the populations were suboptimally adapted.

@@

Finally, despite our model's precise separation of phenotype and kryptotype, this relationship in nature may be far more complicated as aspects of the kryptotype may be less ``hidden'' than we currently assume. For instance, attributes excluded from the phenotype as modelled here, ignore the potential energy costs associated with execessively large (non-minimal) kryptotypes, as well as the relationship between a specific network architecture and mutational robustness, as well as many other presently neglected phenomena. Despite the model's simplifications and potentially nontrivial oversights, it may still roughly reflect and shed light upon speciation patterns as they occur in nature.
